\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-01-31 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

\newcommand{\JGnote}[1]{\fbox{\parbox{\textwidth}{ \color{blue} JG Note $\Rightarrow$ #1}}}
\newcommand{\red}{\textcolor{red}}
\journal{Advances in Water Resources}

@@

\title{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media}

%% use optional labels to link authors explicitly to addresses:
\author[UoA]{K. Christou} \author[UoA,UFRGS]{W.C. Rad\"unz}  \author[UoA]{B. Lashore} 
\author[UoA]{J.L.M.A. Gomes\corref{cor1}}\ead{jefferson.gomes@abdn.ac.uk}

\cortext[cor1]{Corresponding author.}
\address[UoA]{Environmental and Industrial Fluid Mechanics Group, School of Engineering, University of Aberdeen, UK}
\address[UFRGS]{Engineering School, Federal University of Rio Grande do Sul, Brazil}

\begin{abstract} 
Numerical modelling of multiphase flow in porous media has a wide range of applications in science and engineering. Continuity equation and the Darcy's law govern the processes, while pressure, velocity, saturation and permeability are quantities that define the behaviour of the flow. A novel control finite element method (CVFEM) is used to discretize the governing equations that will solve the multi-fluid porous media flow model. The current model is the basis of a general purpose flow simulator that is equipped with anisotropic mesh adaptivity techniques to increase the computational accuracy of numerical methods used in geophysical fluid mechanics. The governing multiphase porous media flow equations are solved using an adaptive and unstructured mesh.
The model is based upon two key numerical characteristics: (a) the family of $P_{n}DG-P_{n+1}$ finite element pairs and (b) a consistent overlapping control volume finite element method (CVFEM) formulation.
Moreover, the mesh adaptivity algorithm employs multi-constraints on element size in different regions of the porous medium to resolve multi-scale transport phenomena. 
%The advantages of mesh adaptivity and the capability of the scheme are demonstrated for simulation of flow in several challenging computational domains.
The scheme captures the key characteristics of the flow while its applicable to heterogeneous porous media and geological formations.
%Numerical modelling of multiphase flows in porous media has application in a wide range of disciplines, from hydrocarbons exploration and extraction to groundwater operations.Viscous fingering, member of the flow instabilities family, is a major concern for the oil \& gas industry, but not only to this. The impact of fingers formation in our simulations is also investigated.  
%This paper describes a  In particular, the \PNDG[1]{1} element (i.e., discontinuous and piecewise linear shape functions to discretise velocity and pressure fields) is introduced as the basis of the discretisation and also as the CVFEM counterpart. In addition to the unstructured grid used for the simulations, uniform and random permeability distribution are introduced. The formulation is applied to saturated porous media systems.
We conclude that understanding the multiphase flow properties will allow us to efficiently facilitate more efficient oil \& gas recovery, efficient contaminant removal and better planning of carbon capture and storage (CCS) projects.

\JGnote{ Abstracts for journal papers have very `squared-structure': (a) Motivation (why would anyone bother to read the paper? Why is it (the paper and/or the subject) relevant? This needs to be covered in 1-3 punchy sentences.); (b) (Main) Objective(s) of the manuscript with particular focus on the \underline{novelty(ies)} the paper may bring (this should be covered in 2-6 (very dry) sentences); (c) Summary of the main technologies that you are using or introducing (and, in some cases, how this/these link(s) with (a), 2-6 sentences); (d) Summary of the main findings (linkage with (b), 2-3 sentences).}

@@

\clearpage
%%%         %%%
%%% SECTION %%%

%%%         %%%
\section{Introduction}\label{section:Intro}

Numerical investigation and modelling of multiphase flows in porous media have attracted the attention of the scientific community over the past 40 years. This is mainly due to the importance of the oil and gas recovery operations, groundwater exploration, underground coal gasification, and more recently on CO$_{\text{2}}$ migration and trapping mechanisms in carbon capture and storage (CCS) operations~\citep{spycher_2003, chen_2006, self_2012, pruess_1990c, white_1981,jiang_2011}.

The conservation of mass along with Darcy's law are the governing equations while quantities such pressure, velocity,saturation and permeability define the  behaviour of the flow. Finite difference methods have been applied to discretise the governing equations \citep{Luo2016,Moortgat2016,Hoteit2008} .The objective is to investigate and accurately describe the behaviour of fluid instabilities in compositional flows using a control volume finite element method (CVFEM). Considering the limitations of finite differences formulations the CVFE The model formulation will use a conservative computational multi-fluid porous media flow model
able to exploit mesh adaptivity methods on a fully-unstructured tetrahedral grid. Also two key characteristics that describe the model above are: (a) the hybrid element pair $P_{n}DGP_{n+1}$ which provides discontinuous and piecewise representation for velocity and pressure fields at FE space and (b) the overlapping control volume finite element method (CVFEM)formulation.

Reservoir engineers and hydro-geologists are dealing with complex geometries, irregular boundaries and heterogeneous geological formations. Forecasting the behaviour of multiphase flows or the future response of the system are major challenges that need to be taken into account. 
%Both permeability and porosity are important parameters in the equations used for numerical simulation of multiphase flow in porous media.\red{(what was the objective of this sentence?)}  %Heterogeniety of permeability and porosity is a product of the medium's formation process %-- deposition and diagenesis -- and exists at different scales. 
To tackle these problems the continuum approach, in a microscopic level, can be applied to a representative elementary Volume (REV) to describe the spatial variation in parameters such as the porosity and permeability~\citep{bastian_phd1999}. %This is well suited for analysis of small scales such as core samples. \red{(These sentences are not linked with the previous sentences, also what is the objective of them? )}.

Larger scale problems can be approached by upscaling techniques. These are techniques which are usually employed to solve problems that fail because of certain types of heterogeneities. These cases are normally resolved by stochastic representation of permeability and porosity fields. 
%Various upscaling techniques, including analytical and numerical, have been introduced in the literature to calculate upscaled flow properties.
There are several approaches to upscale solutions, while three common types are discussed. The $1^{st}$ assumes capillary equilibrium and calculates effective relative permeability using a suitable technique, such as numerical methods, percolation methods and space re-normalization methods. The $2^{nd}$ approach is called pseudoization. It generates a psuedo relative permeability by matching immiscible displacement simulation of a coarse-grid to the output of fine-grid simulations. The $3^{rd}$ approach involves direct calculation of relative permeability from displacement experiments.~\citep{johnson_1959, jones_1978}.

The first approach is limited to low flow rates and very small heterogeneities. The second approach overcomes these limitations but also posses drawbacks such us, unavailability of fine-grid simulation with suitable level of accuracy, dependence of the pseudo functions on rate, direction and grid specification. The experimental data from the third approach is only suitable for a small class of scenarios.~\citep{Chang_1994}. While  non-uniform permeability and porosity may be easy to deal with at a microscopic scale for porous media flow equations, it presents a lot of challenges at a meso/field scale, which is required for reservoir simulations.

\JGnote{As a rule of thumb, when you are introducing a paragraph, the very first sentence needs to give the reader a clue of either linkage with previous paragraphs or linkage with the main subject(s).}

%\medskip
Under certain geological conditions the displacement of fluids in porous media is unstable and results in the formation of irregular flow patterns in the form of fingers %\textit{fingers}. 
The flow instabilities mentioned above are described as \textit{viscous fingering}. Even though the mechanism of the instabilities may vary, 
%(in immiscible fluids the most important factor is the surface tension while in miscible fluids, it is the diffusion \red{(thermal or mass diffusion ...be specific)})
however it is still a function of relative mobility, gravity, dispersion, permeability differences. Moreover, in most cases interfacial tension is directly linked with the mobility ratio between the displaced and the displacing fluid.

@@

%\medskip
%\begin{comment}
Finite difference methods (FDM) have been extensively used in most industry-standard flow simulators~\citep{aziz_1986, chen_2005, chang_1990} with relative success in multi-physics single phase applications (e.g., black oil model, passive tracer advection etc). However, FDM are often limited to relatively simple geometries (with low grid quality and orientation) representing idealised geological formations~\citep{mlacnik_2004, king_1999}. Additionally, FDM produce excessive numerical dispersion when heterogeneity (represented by permeability and porosity fields) is present~\citep{chavent_1986}. The geometrical flexibility associated with high-order numerical accuracy of finite element methods (FEM) has proven to overcome these deficiencies.

@@

@@

In this paper, \textcolor{red}{the main objective is to investigate the impact of heterogeneous porous media in fingers formation using a FEM-based unstructured fixed mesh multi-fluid model with uniform permeability distribution under both saturated and unsaturated porous media}. The overlapping control volume finite element method (OCVFEM), formulation recently introduced by~\citet{jackson_2013}~\citep[see also][]{gomes_2013,pavlidis_2013} is used to investigate the development of multiphase flows through heterogeneous porous media.This finite element based method has a dual pressure-velocity representation in CV and FEM spaces  are embedded into a family of triangular/tetrahedral element pairs, P$_{m}$DG-P$_{n}$~\citep{cotter_2009b} and P$_{m}$DG-P$_{n}DG$~\citep{xie_2014}. The dual velocity and pressure fields are represented by $m^{\rm th}$- (discontinuous) and $({n})^{\rm th}$-order (continuous/discontinuous) polynomials, respectively.
Moreover, velocity field has dual mapping in FEM and CV spaces. This approach can achieve an exact representation of the Darcy force balance equations. 
In addition, the study focused in flow instabilities determined by the viscosity ratio. In the simulations presented in this paper the element pairs as described above and the overlapping control volume finite element method (CVFEM) model were used, in which the dual velocity-pressure fields are embedded in FEM space with simultaneous projection into CV space. Scalar fields (e.g., saturation, concentration, density, etc) are represented in the control volume space. High-order accurate downwind schemes on element boundaries on discontinuous scalar fields are flux-limited (based on NVD approach) to obtain bounded and compressive (capturing the interfaces) solutions. Section $2$ summarises the numerical formulation used in this work, followed by description of the flow instabilities  and a description of model validation. Finally, conclusions are drawn in the final section.

%%%         %%%
%%% SECTION %%%
%%%         %%%
%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Numerical Formulation}\label{section:NumericalFormulation}
In the OCVFEM formulation, a consistent dual pressure-velocity representation in control volume (CV) and finite element (FE) spaces is used with a recently developed family of $P_{m}DG-P_{n}$ tetrahedral element pairs.The new family of $PN_{n}{m}$ element pair was originally introduced. These element-pairs have a discontinuous $m^{th}$-order polynomial representation for velocity and a continuous representation for pressure of order {\it n}.  The mass balance (continuity) equations are solved in CV space and a Petrov-Galerkin finite element (FE) method is used to obtain the high-order fluxes on CV boundaries, which are limited to yield bounded fields. 
This family of element pairs was originally introduced for geophysical fluid dynamics applications. In particular, the $P_{1}DG-P_{2}$ element pair was developed to represent the balance of geostrophic pressure and velocity without introducing spurious pressure modes. This element pair (Fig.~\ref{fig:2}) has a linear discontinuous basis function for velocity $\left(\text{P}_{1}\text{DG}\right)$ and a quadratic polynomial basis function for pressure $\left(\text{P}_{2}\right)$ which maintains $C_{\text{0}}$ continuity across element boundaries. Pressure nodes are located at the element corners and mid-points of the element-edges and control volumes are built around them. Scalar fields, such as saturation, density, temperature, species concentration are stored in control volume space. This family of finite element pairs, when used with the OCVFEM method presented here, is mass conservative and also exactly enforces the force balance represented by the multiphase Darcy equation.

Let's assume a porous media domain $\Omega\subset\mathcal{R}^{n}$ ($n$ is the number of dimensions). Darcy's law for single phase flow can be extended for immiscible multiphase flow as,
\begin{equation}
  \mathbf{q}_{k} = -\frac{\mathcal{K}_{{r}_k}\left(S_{k}\right)\mathbf{K}}{\mu_{k}}\left( \nabla p_{k} - {\mathbf{s}_{u}}_{k} \right),
%S_{k}\left(\lambda_{k}\mathcal{K}\right)^{-1}\phi v_{k} = \underline{\underline{\sigma}}_{k} {\bf u}_{k} = -\nabla p + \mathcal{S}_{u},
\label{Eqn:DarcyEq}
\end{equation}
where $\mathbf{q}_{k}$ is the $k\left(\in\left\{1,N_{p}\right\}\right)$-phase Darcy flow rate. $S$, $\mathbf{K}$, $\mathcal{K}_{rk}\left(S_{k}\right)$ and $\mu$ are saturation, absolute permeability tensor, phase relative permeability and viscosity, respectively. $s_{u}$ is a source term (e.g., gravity force) associated with the force balance and $p$ is the pressure.

>>>>

\newcommand{\JGnote}[1]{\fbox{\parbox{\textwidth}{ \color{red} JG Note $\Rightarrow$ #1}}}
\newcommand{\KCnote}[1]{\fbox{\parbox{\textwidth}{ \color{black} KC Note $\Rightarrow$ #1}}}
\newcommand{\red}{\textcolor{red}}
\newcommand{\blue}{\textcolor{blue}}
\newcommand{\green}{\textcolor{green}}
\journal{Advances in Water Resources}
\newcommand{\frc}{\displaystyle\frac}
\newcommand{\PN}[2][error]{P$_{#1}$DG-P$_{#2}$}
\newcommand{\PNDG}[2][error]{P$_{#1}$DG-P$_{#2}$DG}

@@

\title{ \blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}}

%% use optional labels to link authors explicitly to addresses:
\author[UoA]{K. Christou} \author[UoA,UFRGS]{W.C. Rad\"unz}  \author[UoA]{B. Lashore} \author[UESC]{F.B.S. de Oliveira}
\author[UoA]{J.L.M.A. Gomes\corref{cor1}}\ead{jefferson.gomes@abdn.ac.uk}

\cortext[cor1]{Corresponding author.}
\address[UoA]{Environmental and Industrial Fluid Mechanics Group, School of Engineering, University of Aberdeen, UK}
\address[UFRGS]{Engineering School, Federal University of Rio Grande do Sul, Brazil}
\address[UESC]{Department of Exact and Technological Sciences, State University of Santa Cruz, Bahia, Brazil}

\begin{abstract} 
Numerical modelling of multiphase flow in porous media has a wide range of applications in science and engineering. Viscous fingering, a fluid instability can be a major concern regarding enhanced oil-recovery methods (EOR). Continuity equation and the Darcy's law govern the processes, while pressure, velocity, saturation and permeability are quantities that define the behaviour of the flow. A novel control finite element method (CVFEM) is used to discretize the governing equations that will solve the multi-fluid porous media flow model. The current model is the basis of a general purpose flow simulator that is equipped with anisotropic mesh adaptivity techniques to increase the computational accuracy of numerical methods used in geophysical fluid mechanics and tackle the difficulties that commercial resevoir simulators may rise when using low order finite volume/differences methods based on structured grids. The governing multiphase porous media flow equations are solved using an adaptive and unstructured mesh. The model is based upon two key numerical characteristics: (a) the family of $P_{n}DG-P_{n+1}$ finite element pairs and (b) a consistent overlapping control volume finite element method (CVFEM) formulation.
Moreover, the mesh adaptivity algorithm employs multi-constraints on element size in different regions of the porous medium to resolve multi-scale transport phenomena. The scheme captures the key characteristics of the flow while its applicable to heterogeneous porous media and geological formations. We conclude that understanding the multiphase flow properties will allow us to efficiently facilitate more efficient oil \& gas recovery, efficient contaminant removal and better planning of carbon capture and storage (CCS) projects.

%\JGnote{ Abstracts for journal papers have very `squared-structure': (a) Motivation (why would anyone bother to read the paper? Why is it (the paper and/or the subject) relevant? This needs to be covered in 1-3 punchy sentences.); (b) (Main) Objective(s) of the manuscript with particular focus on the \underline{novelty(ies)} the paper may bring (this should be covered in 2-6 (very dry) sentences); (c) Summary of the main technologies that you are using or introducing (and, in some cases, how this/these link(s) with (a), 2-6 sentences); (d) Summary of the main findings (linkage with (b), 2-3 sentences).}

@@

\clearpage
%%%         %%%
%%% SECTION %%%
%%%         %%%

\KCnote{Please reorganise the structure like the papers you read, and reference them!}

\section{Introduction}\label{section:Intro}

Numerical investigation and modelling of multiphase flows in porous media have attracted the attention of the scientific community over the past 40 years. Characterization of the flow of fluids through porous media serves as the foundation of reservoir engineering studies. Reservoir engineers and hydro-geologists are dealing with complex geometries, irregular boundaries and heterogeneous geological formations. Forecasting the behaviour of multiphase flows or the future response of the system are major challenges that need to be taken into account.  This is mainly due to the importance of the oil and gas recovery operations, groundwater exploration, underground coal gasification, and more recently on CO$_{\text{2}}$ migration and trapping mechanisms in carbon capture and storage (CCS) operations~\citep{spycher_2003, chen_2006, self_2012, pruess_1990c, white_1981,jiang_2011}.

\medskip

The conservation of mass along with Darcy's law are the governing equations that describe the flow pattern. Particle methods, front-tracking methods \citep{univerdi_1992} but in particular volume of fluid (VOF) methods \citep{hirt_1981,rider_1998} are used to describe the afformentioned proeprties. In addition, finite difference methods have been applied to discretise the governing equations \citep{Luo2016,Moortgat2016,Hoteit2008} .

The objective of this work is to investigate and accurately describe the behaviour of fluid instabilities in compositional flows using a control volume finite element method (CVFEM). Considering the limitations of finite differences formulations the CVFE The model formulation will use a conservative computational multi-fluid porous media flow model able to exploit mesh adaptivity methods on a fully-unstructured tetrahedral grid. Also two key characteristics that describe the model above are: (a) the hybrid element pair $P_{n}DGP_{n+1}$ which provides discontinuous and piecewise representation for velocity and pressure fields at FE space and (b) the overlapping control volume finite element method (CVFEM)formulation.

Larger scale problems can be approached by upscaling techniques. These are techniques which are usually employed to solve problems that fail because of certain types of heterogeneities. These cases are normally resolved by stochastic representation of permeability and porosity fields. There are several approaches to upscale solutions, while three common types are discussed. The $1^{st}$ assumes capillary equilibrium and calculates effective relative permeability using a suitable technique, such as numerical methods, percolation methods and space re-normalization methods. The $2^{nd}$ approach is called pseudoization. It generates a psuedo relative permeability by matching immiscible displacement simulation of a coarse-grid to the output of fine-grid simulations. The $3^{rd}$ approach involves direct calculation of relative permeability from displacement experiments.~\citep{johnson_1959, jones_1978}. Problems start to rise when complex flow patterns will appear during the recovery of hydrocarbons triggered by heterogeneities.  
%The first approach is limited to low flow rates and very small heterogeneities. The second approach overcomes these limitations but also posses drawbacks such us, unavailability of fine-grid simulation with suitable level of accuracy, dependence of the pseudo functions on rate, direction and grid specification. The experimental data from the third approach is only suitable for a small class of scenarios.~\citep{Chang_1994}. While  non-uniform permeability and porosity may be easy to deal with at a microscopic scale for porous media flow equations, it presents a lot of challenges at a meso/field scale, which is required for reservoir simulations.

%\JGnote{As a rule of thumb, when you are introducing a paragraph, the very first sentence needs to give the reader a clue of either linkage with previous paragraphs or linkage with the main subject(s).}

\medskip
Under certain geological conditions the displacement of fluids in porous media is unstable and results in the formation of irregular flow patterns in the form of fingers %\textit{fingers}. 
These patterns are also described as flow instabilities or \textit{viscous fingering},  or most widely known as Saffman-Taylor instabilitises. This phenomenon will occur during an unfavorable -mobility-ration displacement, when a more viscous fluid is displaced by a less viscous (Saffman and Taylor $1958$). The mechanism of the instabilities may vary, 
%(in immiscible fluids the most important factor is the surface tension while in miscible fluids, it is the diffusion \red{(thermal or mass diffusion ...be specific)})
however it is still a function of relative mobility, gravity, viscous forces and permeability differences. In conventional reservoirs the principal forces that affect fluid flow include, viscous forces, effects of gravity and permeability diffrences while in unconventional gas reservoirs flow simulation models take into account the phenomena of adsorption and diffusion \citep{abdus_2015}.
Fingering will occur in miscible displacments when the viscocity ratio is greater than unity. As surface tension becomes weak, the front of the steady finger is susceptible to a viscous-fingering instability by the basic mechanism associated with a less viscous fluid displacing a more viscous one. After a split, each of the new lobes of the finger is stable as a result of their being thinner than the finger from which they split. As a result of \textit{shielding}, one of these lobes will eventually outgrow the others and will then spread to occupy the appropriate width of the cell. In the process, it reaches a width that is again unstable to splitting, and the pattern repeats. Thus, surface tension plays an essential dual role; it must be weak enough for the tip front to be unstable, but it is also the physical force causing the \textit{spreading} and ensuing repeated branching. Important parameter is the mobility (or viscosity) ratio. This is given by Eqn.~\ref{eq:mobility-of-fluid} where $K$ is absolute permeability and $k_{r}$ is relative permeability :

\begin{equation}
\lambda = \frac{K\;k_{\alpha}}{\mu}
\label{eq:mobility-of-fluid}
\end{equation}

@@

% are based on domains that require high resolution static meshes to capture the development of the fingers 
Even though some commercial simulators provide the user with option of an unstructured type of mesh still they are not able to dynamically adapt the mesh as the simulation develops and they increase the computational cost as the saturation front changes slowly. In this paper the use of dynamic mesh adaptivity to capture the viscous fingering in reservoir flows is introduced, as one of the methods that will allow the user to focus the effort where is needed, minimizing both the CPU time and the numerical diffusion.

The remainder of the paper is orgranised as follows.

\KCnote{ this needs to be rephrased later -- Next section \citet{math-model} illustrates the model and the governing equations of immiscible and incompressible flow in porous media. The method and numerical formulations that describes the CVFEM follows, along with cases that demonstrate how the mesh adaptivity for simulation of viscous fingering is implemented. In the end, conclusions and final remarks of the work are discussed.--}

\section{Mathematical Formulation}\label{section:math-model}
%\medskip
%\begin{comment}
The displacing and displaced fluids are considered to be incompressible and immiscible under isothermal and no source or sink conditions. In addition, we neglect the effects of capillarity and gravity in our simulations. Therefore the governing equations of conservation of mass and Darcy's law can be described below as,

\begin{equation}\label{e:darcy_eqn}
  \mathbf{q}_{\alpha} = -\frac{\mathcal{K}_{{r}_\alpha}\mathbf{K}}{\mu_{\alpha}}\left(\nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\mathbf{q}_{\alpha}$ is the $\alpha$-th phase Darcy velocity, $\mathbf{K}$ is the absolute permeability tensor of the porous medium, $\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ is the phase relative permeability, which is a function of the phase saturation $S_{\alpha}\left(\mathbf{r},t\right)$. $\mu_{\alpha}$, $p_{\alpha}$, $\rho_{\alpha}$, and $\mathbf{s}_{{u}_\alpha}$ are the phase dynamic viscosity, pressure, density and source term, which may include gravity and/or capillarity, respectively.

Introducing a saturation-weighted Darcy velocity defined as $\mathbf{u}_\alpha= \mathbf{q}_\alpha/S_\alpha$, then Eqn.~\ref{e:darcy_eqn} may be rewritten as:
\begin{equation}
  \mathbf{v}_\alpha={\underline {\underline \sigma}}_{\alpha} \mathbf{u}_{\alpha} = - \nabla p_{\alpha} + {\mathbf{s}_{u}}_{\alpha},
  \label{force-bal}
\end{equation}

where ${\underline {\underline \sigma}}_{\alpha}=\mu_\alpha S_\alpha \left(\mathcal{K}_{{r}_\alpha}\mathbf{K}\right)^{-1}$ represents the implicit linearisation of the viscous frictional forces. Assuming a porous media domain $\Omega\subset\mathcal{R}^{n}$ ($n$ is the number of dimensions). Darcy's law for single phase flow can be extended (from Eq.\ref{e:darcy_eqn}) for immiscible multiphase flow to Eq.\ref{E:darcy_eqn} below,

\begin{equation}
  \mathbf{q}_{\alpha} = -\frac{\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)\mathbf{K}}{\mu_{\alpha}}\left( \nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
%S_{k}\left(\lambda_{k}\mathcal{K}\right)^{-1}\phi v_{\alpha} = \underline{\underline{\sigma}}_{\alpha} {\bf u}_{\alpha} = -\nabla p + \mathcal{S}_{u},
\label{E:darcy_eqn}
\end{equation}
where $\mathbf{q}_{\alpha}$ is the $k\left(\in\left\{1,N_{p}\right\}\right)$-phase Darcy flow rate. $S$, $\mathbf{K}$, $\mathcal{K}_{r\alpha}\left(S_{\alpha}\right)$ and $\mu$ are saturation, absolute permeability tensor, phase relative permeability and viscosity, respectively. $s_{u}$ is a source term (e.g., gravity force) associated with the force balance and $p$ is the pressure.

 Defining the advective velocity averaged over the entire domain -- $\mathbf{u}_{\alpha}= \mathbf{q}_{\alpha}/S_{\alpha}$, then we may rewrite Eq.(\ref{E:darcy_ext}),
\begin{equation}
  {\underline {\underline \sigma}}_{\alpha} \mathbf{u}_{\alpha} = - \nabla p + {\mathbf{s}_{u}}_{\alpha},
  \label{E:darcy_ext}
\end{equation}
$\underline{\underline{\sigma}}_{\alpha}=\mu_{\alpha}S_{\alpha}\left(\mathcal{K}_{r\alpha}\mathbf{K}\right)^{-1}$ represents the implicit linearisation of the viscous frictional forces, and is piecewise constant within each FE and is obtained via basis functions local to each CV within each element \citep{Jackson13,Radunz14}.

\begin{eqnarray}
\left(\displaystyle\frac{\mu_{\alpha}S_{\alpha}}{{\mathbf K}\mathcal{K}_{r\alpha}}\right) {\mathbf u}_{\alpha} = \underline{\underline{\sigma}}_{\alpha} {\mathbf u}_{\alpha} = -\nabla p + \mathcal{S}_{u,\alpha},\;\;\; x_{i}\in\Omega, t>0 \label{eqn:darcy_eqn} \\
\phi\displaystyle\frac{\partial S_{\alpha} }{\partial t} +   \nabla \cdot \left( {\mathbf u}_{\alpha}  S_{\alpha}\right) =  \mathcal{S}_{cty,\alpha},\;\;\; x_{i}\in\Omega, t>0\label{eqn:saturation_eqn}
\end{eqnarray}

where $S$, $\mu$, ${\bf K}$, $p$ and $\phi$ are the saturation, viscosity, absolute permeability, pressure and porosity, respectively. ${\mathbf u}_{k}$ is the saturation-weighted Darcy velocity of the $k$-phase and $\mathcal{K}_{r,\alpha}$ is the relative permeability. $\mathcal{S}$ is the source term associated with the Darcy and continuity equations. $S_{\alpha}$ is saturation of the $\alpha$-phase with the constraint of $\sum\limits_{\alpha=1}^{\mathcal{N}_{p}} S_{\alpha} = 1$, where $\mathcal{N}_{p}$ denotes the number of phases.

\medskip
Finite element basis functions for velocity and pressure fields are introduced in the discretisation of force-balance equation -- Eqn.~\ref{eqn:darcy_eqn}. Hybrid (i.e., overlapping) basis functions are also developed to allow CV-based velocity to be extrapolated across the entire element. $\underline{\underline{\sigma}}_{k}=\underline{\underline{\sigma}}_{k}\left({\mathbf K}, \mathcal{K}_{rk}, S_{K}, \mu_{k}\right)$ is an absorption-like term that lies in both CV and FEM spaces. Though, whilst the saturation field is calculated in CV space, permeability $\left({\mathbf K}\right)$ is piecewise constant in FE space to allow efficient representation of surface parametrised geometries. The global saturation equation (Eqn.~\ref{eqn:saturation_eqn}) is discretised in space with CV basis function and with the $\theta$-method in time as described by \citet{gomes_book_2012}. Velocities across CV interfaces (within and between elements, see fig.\ref{fig:fem_cv} below) are calculated through a directional-weighted flux-limited scheme based on upwind value of $\sigma$ at individual CV as described by \citet{gomes_2013}~\citep[see also][]{jackson_2013}.

\begin{figure}[h]
\centering
\includegraphics[width=.5\textwidth]{./Pics/P1DGP2.PNG}
\caption{2D representation of the element pairs presented in this work. Shaded areas denote control volumes (in which saturation is stored), black points represent the pressure nodes and white points the velocity nodes. Note that in (b) velocity and pressure nodes overlap in the triangles' vertices.}
\label{fig:fem_cv}
\end{figure}

\section{The Overlapping CVFEM formulation}

Finite difference methods (FDM) have been extensively used in most industry-standard flow simulators~\citep{aziz_1986, chen_2005, chang_1990} with relative success in multi-physics single phase applications (e.g., black oil model, passive tracer advection etc). Analytic models can be used to analyze reservoir performance as long as the reservoir is not highly heterogeneous. However, FDM are often limited to relatively simple geometries 
%(with low grid quality and orientation) 
representing idealised geological formations~\citep{mlacnik_2004, king_1999}. Additionally, FDM produce excessive numerical dispersion when heterogeneity (represented by permeability and porosity fields) is present~\citep{chavent_1986}. The geometrical flexibility associated with high-order numerical accuracy of finite element methods (FEM) has proven to overcome these deficiencies.

@@

In the CVFE method pressure is interpollated using piecewise linear finite element basis functions (Eq.\ref{cvfe_pressure}) while saturation is represented in node centered CV basis (Eq.\ref{cvfe_saturation}) control volume basis functions respectevely:

\begin{equation}
 p^{n} = \sum_{j} \hat{p_{j}} N_{\alpha} 
\label{cvfe_pressure}
\end{equation}

\begin{equation}
 S^{n} = \sum_{j} \hat{p_{j}} M_{b} 
\label{cvfe_saturation}
\end{equation}

where $N_{\alpha}$ is first order finite element basis function and $M_{b}$ is the control volume basis function that is equal to unity over the control volume constructed around node j and is equal to zero everywhere else. The saturation equation is solved explicitly after solving for pressure at each time.

@@

%\textcolor{red}{The impact of heterogeneous porous media in fingers formation using a FEM-based unstructured fixed mesh multi-fluid model with uniform permeability distribution under both saturated and unsaturated porous media}. The overlapping control volume finite element method (OCVFEM), formulation recently introduced by~\citet{jackson_2013}~\citep[see also][]{gomes_2013,pavlidis_2013} is used to investigate the development of multiphase flows through heterogeneous porous media.This finite element based method has a dual pressure-velocity representation in CV and FEM spaces  are embedded into a family of triangular/tetrahedral element pairs, P$_{m}$DG-P$_{n}$~\citep{cotter_2009b} and P$_{m}$DG-P$_{n}DG$~\citep{xie_2014}. The dual velocity and pressure fields are represented by $m^{\rm th}$- (discontinuous) and $({n})^{\rm th}$-order (continuous/discontinuous) polynomials, respectively.
%Moreover, velocity field has dual mapping in FEM and CV spaces. This approach can achieve an exact representation of the Darcy force balance equations. 
%In addition, the study focused in flow instabilities determined by the viscosity ratio. In the simulations presented in this paper the element pairs as described above and the overlapping control volume finite element method (CVFEM) model were used, in which the dual velocity-pressure fields are embedded in FEM space with simultaneous projection into CV space. Scalar fields (e.g., saturation, concentration, density, etc) are represented in the control volume space. High-order accurate downwind schemes on element boundaries on discontinuous scalar fields are flux-limited (based on NVD approach) to obtain bounded and compressive (capturing the interfaces) solutions. Section $2$ summarises the numerical formulation used in this work, followed by description of the flow instabilities  and a description of model validation. Finally, conclusions are drawn in the final section.

%Numerical investigation and modelling of multiphase flows in porous media have attracted the attention of the scientific community over the past 40 years. Reservoir engineers and hydro-geologists are dealing with complex geometries, irregular boundaries and heterogeneous geological formations. Forecasting the behaviour of multiphase flows or the future response of the system are major challenges that need to be taken into account. This is mainly due to the importance of the oil and gas recovery operations, groundwater exploration, underground coal gasification, and more recently on CO$_{\text{2}}$ migration and trapping mechanisms in carbon capture and storage (CCS) operations~\citep{spycher_2003, chen_2006, self_2012, pruess_1990c, white_1981,jiang_2011}.

The conservation of mass along with Darcy's law are the governing equations while variables such as pressure, velocity,saturation and permeability define the behaviour of the flow. Finite difference methods have been applied to discretise the governing equations \citep{Luo2016,Moortgat2016,Hoteit2008} .The objective of this work is to investigate and accurately describe the behaviour of fluid instabilities in compositional flows using a control volume finite element method (CVFEM). Most commercial reservoir simulators use low-order finite-volume/differences methods based on structured grids. Considering the limitations of finite differences formulations, the CVFE The model formulation will use a conservative computational multi-fluid porous media flow model able to exploit mesh adaptivity methods on a fully-unstructured tetrahedral grid. Also two key characteristics that describe the model above are: (a) the hybrid element pair $P_{n}DGP_{n+1}$ which provides discontinuous and piecewise representation for velocity and pressure fields at FE space and (b) the overlapping control volume finite element method (CVFEM)formulation.

%Both permeability and porosity are important parameters in the equations used for numerical simulation of multiphase flow in porous media.\red{(what was the objective of this sentence?)}  %Heterogeniety of permeability and porosity is a product of the medium's formation process %-- deposition and diagenesis -- and exists at different scales. 
%To tackle these problems the continuum approach, in a microscopic level, can be applied to a representative elementary Volume (REV) to describe the spatial variation in parameters such as the porosity and permeability~\citep{bastian_phd1999}. %This is well suited for analysis of small scales such as core samples. \red{(These sentences are not linked with the previous sentences, also what is the objective of them? )}.

%Larger scale problems can be approached by upscaling techniques. These are techniques which are usually employed to solve problems that fail because of certain types of heterogeneities. These cases are normally resolved by stochastic representation of permeability and porosity fields. 
%Various upscaling techniques, including analytical and numerical, have been introduced in the literature to calculate upscaled flow properties.
%There are several approaches to upscale solutions, while three common types are discussed. The $1^{st}$ assumes capillary equilibrium and calculates effective relative permeability using a suitable technique, such as numerical methods, percolation methods and space re-normalization methods. The $2^{nd}$ approach is called pseudoization. It generates a psuedo relative permeability by matching immiscible displacement simulation of a coarse-grid to the output of fine-grid simulations. The $3^{rd}$ approach involves direct calculation of relative permeability from displacement experiments.~\citep{johnson_1959, jones_1978}.

%The first approach is limited to low flow rates and very small heterogeneities. The second approach overcomes these limitations but also posses drawbacks such us, unavailability of fine-grid simulation with suitable level of accuracy, dependence of the pseudo functions on rate, direction and grid specification. The experimental data from the third approach is only suitable for a small class of scenarios.~\citep{Chang_1994}. While  non-uniform permeability and porosity may be easy to deal with at a microscopic scale for porous media flow equations, it presents a lot of challenges at a meso/field scale, which is required for reservoir simulations.
